---
title: "651 final project"
output: html_notebook
---


```{r}
kidney <- read.csv("pseudo_kidney_transplant_2005.csv")

```

# Models

### Effect of gender on survival (are male-male/female-female grafts more successful?)
```{r}
# create an indicator variable for same sex of donor and recipient
kidney$same_sex_don_rec <- ifelse(kidney$DON_GENDER==kidney$REC_GENDER, 1,0)
kidney$success <- ifelse(kidney$event==1,1,0) # we are defining success as survival
kidney$graft_fail <- ifelse(kidney$event==3,1,0)
kidney$death <- ifelse(kidney$event==2,1,0)

gender_glm_success <- glm(data=kidney, success ~  same_sex_don_rec, family = binomial())
summary(gender_glm_success)

gender_glm_graft_fail <- glm(data=kidney, graft_fail ~  same_sex_don_rec, family = binomial())
summary(gender_glm_graft_fail)

gender_glm_death <- glm(data=kidney, death ~  same_sex_don_rec, family = binomial())
summary(gender_glm_death)
```

Not technically significant on survival but will definitely need to control for things like age and weight for this model. Does not appear to have significant effect on graft success or death.

### Effect of age on recipient survival
```{r}
age_glm <- glm(data=kidney, success ~  DON_AGE + REC_AGE_AT_TX, family = binomial())
summary(age_glm)

age_glm_graft_fail <- glm(data=kidney, graft_fail ~  DON_AGE + REC_AGE_AT_TX, family = binomial())
summary(age_glm_graft_fail)

age_glm_death <- glm(data=kidney, death ~  DON_AGE + REC_AGE_AT_TX, family = binomial())
summary(age_glm_death)
```

It looks like donor age definitely has a significant effect on survival.
Donor age and age of recipient at time of transplant both have significant effects on graft failure
Age of recipient at time of transplant has a significant effect on death




# Reduced logistic model
Living v. death ~ age of donor + age of recipient + recipient diab + recipient bmi + donor hyp + donor diab + Donor hypertension*donor diabetes + Recipient diabetes*recipient bmi + same race
```{r}
kidney$success <- ifelse(kidney$event==1,1,0) # we are defining success as survival, graft failure and death are grouped together
kidney$same_race <- ifelse(kidney$DON_RACE==kidney$REC_RACE,1,0)
kidney$rec_diab_yn <- ifelse(kidney$REC_DIAB=="None",0,1)
# make age groups
#Donors: under 18, 18-39, 40-65, 65+
#Recipients: 18-39, 40-65, 65+
kidney$DON_AGE_CAT <- cut(kidney$DON_AGE, 
                          breaks = c(0,18,39,65,80),
                          labels = c("under 18","18-39","40-65","65+"))
kidney$REC_AGE_AT_TX_CAT <- cut(kidney$REC_AGE_AT_TX,
                                breaks = c(18,39,65,90),
                                labels = c("18-39","40-65","65+"))

kidney_2 <- na.omit(kidney) # omit patients with NA's in any column
glm.red <- glm(formula = success ~ DON_AGE_CAT + REC_AGE_AT_TX_CAT + REC_DIAB + REC_BMI +
                 DON_HTN + DON_HIST_DIAB + DON_HTN*DON_HIST_DIAB + 
                 REC_DIAB*REC_BMI + same_race + REC_AGE_AT_TX_CAT*DON_AGE_CAT,
                 data = kidney,
                 family = binomial(link=logit))
glm.red_noagecat <- glm(formula = success ~ DON_AGE + REC_AGE_AT_TX + REC_DIAB + REC_BMI +
                 DON_HTN + DON_HIST_DIAB + DON_HTN*DON_HIST_DIAB + 
                 REC_DIAB*REC_BMI + same_race + REC_AGE_AT_TX*DON_AGE,
                 data = kidney,
                 family = binomial(link=logit))

summary(glm.red)
#summary(glm.red_noagecat)

# Deviance/Pearson's goodness of fit test
D = 2*(2981.0-2862.3)

```

Deviance goodness of fit test: Testing the null hypothesis that the model with the intercept only is better against the alternative hypothesis that the model with all the covariates is the better model. D = 2(2981.0-2862.3) = 237.4 is ~$\chi^2_{22}$. The critical point for $\chi^2_{22}$ is 33.924. Since 237.4 > 33.924, we reject the null hypothesis and conclude that the model with all our covariates is the better model.



Should we also be looking at specific type of diabetes? Type 1 vs type 2?


# Multinomial Regression

We may have to remove the patients who have an NA in any of their columns, but it isnt that many, its only 19
```{r}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(knitr, gmodels, car, lmtest, VGAM, nnet)

kidney_2 <- na.omit(kidney) # omit patients with NA's in any column
kidney_2$REC_BMI_c <- kidney_2$REC_BMI-mean(kidney_2$REC_BMI)# center the bmi for interpretation purposes
multi_mod_full <- vglm(as.factor(event) ~ DON_AGE_CAT + REC_AGE_AT_TX_CAT + rec_diab_yn + REC_BMI +
                 DON_HTN + DON_HIST_DIAB + DON_HTN*DON_HIST_DIAB + 
                 rec_diab_yn*REC_BMI + same_race + REC_AGE_AT_TX_CAT*DON_AGE_CAT,
                 data = kidney_2,
                 family = multinomial(refLevel=1))

summary(multi_mod_full)

multi_mod_red <- vglm(as.factor(event) ~ DON_AGE_CAT + REC_AGE_AT_TX_CAT + rec_diab_yn + REC_BMI +
                 DON_HTN + DON_HIST_DIAB + DON_HTN*DON_HIST_DIAB + 
                 rec_diab_yn*REC_BMI + same_race,
                 data = kidney_2,
                 family = multinomial(refLevel=1))

summary(multi_mod_red)
#data.frame(c(coef(multi_mod), as.vector(sqrt(diag(vcov(multi_mod))))[3:4]))

# Can't do deviance test so doing hosmer-lemeshow for multinomial regression
#install.packages("generalhoslem")
#library(generalhoslem)
logitgof(kidney_2$event, fitted(multi_mod_full))
logitgof(kidney_2$event, fitted(multi_mod_red))
```
For the Hosmer-Lemeshow test we are testing the null hypothesis $\H_0:$ the model fits the data well against the alternative hypothesis $H_1:$ the model does not fit the data well. We have a p-value of 0.047 for this test so we reject the null and conclude that the model does not fit the data well.



## Proportional odds multinomial regression
```{r}
kidney$ord_event<-factor(kidney$event, levels=c("1", "2", "3"), ordered = T) # We would need to specify the order for response variable
propOdd.vglm<-vglm(ord_event~ DON_AGE + REC_AGE_AT_TX + rec_diab_yn + REC_BMI +
                 DON_HTN + DON_HIST_DIAB + DON_HTN*DON_HIST_DIAB + 
                 rec_diab_yn*REC_BMI + same_race ,data=kidney, family = cumulative(link = "logitlink", parallel = TRUE))

summary(propOdd.vglm)
NoPropOdd.vglm<-vglm(ord_event~ DON_AGE + REC_AGE_AT_TX + rec_diab_yn + REC_BMI +
                 DON_HTN + DON_HIST_DIAB + DON_HTN*DON_HIST_DIAB + 
                 rec_diab_yn*REC_BMI + same_race ,data=kidney, family = cumulative(link = "logitlink", parallel = FALSE))
summary(NoPropOdd.vglm)
# To fit the model without proportional odds assumption, we could just set parallel argument as false
lrtest(NoPropOdd.vglm,propOdd.vglm)
prop_test<-lrtest(NoPropOdd.vglm,propOdd.vglm)
```






